### 2010-12-28 Documentation/atropos/atropos.tex
ENH:  A few minor edits.
git-svn-id: https://advants.svn.sourceforge.net/svnroot/advants@719 fd2da0a4-d968-48df-9cb5-d8c496287cbf

\subsection{Initializing the Atropos Objective}
Atropos has a number of parameters defined within
Listing \ref{listing:n4} and will function on 2, 3 or 4 dimensional
data.  However, the majority of the time, users will be concerned with
a smaller set of input parameters.  Here, we list the recommended input and
an example definition for each parameter:
\begin{description}
\item[Input images to be segmented:] If more than
  one input image is passed, then a multivariate model will be
  instantiated.  E.g.  {\ttfamily -a Image.nii.gz} for one image and
  {\ttfamily -a Image1.nii.gz -a Image2.nii.gz} for multiple images.
\item[Input image mask:] This binary image
  defines the spatial segmentation domain.  Voxels outside the
  masked region are designated with the label $0$.  E.g.  {\ttfamily -x mask.nii.gz}. 
\item[Convergence criteria:] The algorithm terminates if it reaches  the maximum number of iterations or
  produces a change less than the minimum threshold change in the posterior.
  E.g. {\ttfamily -c [5,1.e-5]}. 
\item[MRF prior:] The key parameter to increase or decrease the
  spatial smoothness of the label map is $\beta$.  A useful range of
  $\beta$ values is $0$ to $0.5$ where we usually use $0.05, 0.1, 0.2$
  in brain segmentation.  E.g.  {\ttfamily -m [ 0.1, 1x1x1 ]} would
  define $\beta=0.1$ with a MRF radius of one voxel in each of three dimensions.
\item[Initialization:] The initialization options include (where the
  first parameter defines $K$, here 3 for each below),
\begin{itemize}
\item  {\ttfamily -i  Kmeans[3]} standard K-means initialization for
  three classes,
\item  {\ttfamily -i  PriorLabelImage[3,label\_image.nii.gz]} and
%  where scalar $w \in [0,1]$ is unused. 
\item  {\ttfamily -i  PriorProbabilityImages[3,label\_prob\%02d.nii.gz,w]} 
where $w=0$~(use the prior probability images only for
  initialization) or $w>0.0$ (use the prior probability images
  throughout the optimization).  If one chooses $0 < w < 1.0$ then one
  will increase (from zero) the weight on the priors.  These images,
  like the {\ttfamily PriorLabelImage}, should be defined with the same domain as
  the input images to be segmented.  
\end{itemize}
\item[Posterior formulation:]  The user may choose to estimate the
  mixture proportions (or not) by setting  {\ttfamily -p Socrates[1]}
  or {\ttfamily -p Socrates[0] }.  Fixed label boundary conditions may be employed by
  selecting the {\ttfamily PriorLabelImage} initialization and   {\ttfamily -p Plato[0]}.
\item[Output:] Atropos will output the hard segmentation and the
  probability image for each model.  E.g.  {\ttfamily -o
    [segmentation.nii.gz,seg\_prob\%02d.nii.gz]} will write
  out the hard segmentation in the first output parameter and a
  probability image for each class named, here, seg\_prob01.nii.gz,
  seg\_prob02.nii.gz, etc. 
\end{description}
Higher dimensions than 4 are possible although we have not yet encountered such 
application-specific need.  Multiple images (assumed to be of the same
dimension, size, origin, etc.), will automatically enable multivariate
likelihoods.  In that case, the first image specified on the 
command line is used to initialize the Random, Otsu, or K-means labeling with the latter initialization further
refined by incorporating the additional intensity images, i.e. an initial univariate K-means clustering is determined 
from the first intensity image which, along with the other images, provides the starting multivariate cluster centers for a 
follow-up multivariate K-means labeling.    More details on each of
the key implementation options are given below.

>>>>

\subsection{Initializing the Atropos Objective}
Atropos has a number of parameters defined within
Listing \ref{listing:n4} and will function on 2, 3 or 4 dimensional
data.  However, the majority of the time, users will be concerned with
a smaller set of input parameters.  Here, we list the recommended input and
an example definition for each parameter:
\begin{description}
\item[Input images to be segmented:] If more than
  one input image is passed, then a multivariate model will be
  instantiated.  E.g.  {\ttfamily -a Image.nii.gz} for one image and
  {\ttfamily -a Image1.nii.gz -a Image2.nii.gz} for multiple images.
\item[Input image mask:] This binary image
  defines the spatial segmentation domain.  Voxels outside the
  masked region are designated with the label $0$.  E.g.  {\ttfamily -x mask.nii.gz}. 
\item[Convergence criteria:] The algorithm terminates if it reaches  the maximum number of iterations or
  produces a change less than the minimum threshold change in the posterior.
  E.g. {\ttfamily -c [5,1.e-5]}. 
\item[MRF prior:] The key parameter to increase or decrease the
  spatial smoothness of the label map is $\beta$.  A useful range of
  $\beta$ values is $0$ to $0.5$ where we usually use 0.05, 0.1 or 0.2
  in brain segmentation.  E.g.  {\ttfamily -m [0.1, 1x1x1]} would
  define $\beta=0.1$ with a MRF radius of one voxel in each of three dimensions.
\item[Initialization:] The initialization options include (where the
  first parameter defines $K$, here 3 for each below),
\begin{itemize}
\item  {\ttfamily -i  Kmeans[3]} standard K-means initialization for
  three classes,
\item  {\ttfamily -i  PriorLabelImage[3,label\_image.nii.gz]} and
%  where scalar $w \in [0,1]$ is unused. 
\item  {\ttfamily -i  PriorProbabilityImages[3,label\_prob\%02d.nii.gz,w]} 
where $w=0$~(use the prior probability images only for
  initialization) or $w>0.0$ (use the prior probability images
  throughout the optimization).  If one chooses $0 < w < 1.0$ then one
  will increase (from zero) the weight on the priors.  These images,
  like the {\ttfamily PriorLabelImage}, should be defined with the same domain as
  the input images to be segmented.  
\end{itemize}
\item[Posterior formulation:]  The user may choose to estimate the
  mixture proportions (or not) by setting  {\ttfamily -p Socrates[1]}
  or {\ttfamily -p Socrates[0] }.  Fixed label boundary conditions may be employed by
  selecting the {\ttfamily PriorLabelImage} initialization and   {\ttfamily -p Plato[0]}.
\item[Output:] Atropos will output the hard segmentation and the
  probability image for each model.  E.g.  {\ttfamily -o
    [segmentation.nii.gz,seg\_prob\%02d.nii.gz]} will write
  out the hard segmentation in the first output parameter and a
  probability image for each class named, here, seg\_prob01.nii.gz,
  seg\_prob02.nii.gz, etc. 
\end{description}
Higher dimensions than 4 are possible although we have not yet encountered such 
application-specific need.  Multiple images (assumed to be of the same
dimension, size, origin, etc.), will automatically enable multivariate
likelihoods.  In that case, the first image specified on the 
command line is used to initialize the Random, Otsu, or K-means labeling with the latter initialization further
refined by incorporating the additional intensity images, i.e. an initial univariate K-means clustering is determined 
from the first intensity image which, along with the other images, provides the starting multivariate cluster centers for a 
follow-up multivariate K-means labeling.    More details on each of
the key implementation options are given below.
